Spatio-temporal estimates of HIV risk group proportions for adolescent girls and young women across 13 priority countries in sub-Saharan Africa

### 2022-04-19 src/docs_12-05-22-mrc-gida/12-05-22-mrc-gida.Rmd
Minimal viable version of these slides #112

1. Enable countries to implement prevention approaches stratified by incidence and behaviour
2. Assess the extent to which this would be beneficial

@@

1. Estimate the proportion of AGYW in four sexual risk behaviour groups at a district level (in 13 countries identified as priority by The Global Fund)
2. Analyze the number of new infections reached by stratified prevention strategies

@@

* We used sexual behaviour data from household surveys (see next slide)
* District-level HIV incidence estimates from the Naomi model [@eaton2021naomi]
* District boundaries

@@

### Stage 1.

* Multinomial logistic regression model for the proportion of AGYW in the None, Low and {High, Very High} risk groups, using all 47 surveys
* Selected model included:
  * Age effects (IID)
  * Country effects (IID)
  * Correlated spatial effects (ICAR)
  * Correlated temporal effects (AR1) 
* Multinomial-Poisson transformation allowed use of `R-INLA` for inference

@@

### Stage 2.

* Logistic regression model for the proportion of {High, Very High} who are in the Very High risk group, using only the 13 surveys with a specific transactional sex question
* Selected model included:
  * Age effects (IID)
  * Country effects (IID)
  * Correlated spatial effects (ICAR)
  * Clients of FSW covariates
* Used `R-INLA` for inference

@@

* Take 1000 samples from each model, then suitably multiply to generate estimates for all four risk groups
* We adjusted the Very High risk group to match national FSW estimates from @johnston2022deriving

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ estimates of risk group proportions $\rho_{itak}$ by district, year and age group
\end{tcolorbox}
\end{center}

@@

```{r echo=FALSE, out.width = "400px"}
knitr::include_graphics("depends/3p1-within-between-country-variation.png")
```

@@

1. Estimates even where there isn't direct data
    * Although some people think of this as "making up data", the data almost never "speak for themselves"
    * Uncertainty should be higher in regions with infilling: important to clearly communicate this
2. Borrow information across space, between countries, over multiple surveys, (weakly) between risk groups
    * See next slide for illustration of why this is important

@@

@@

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ estimates of HIV incidence $\lambda_{iak}$ and number of new HIV infections $I_{iak}$ by district, age group and risk group
\end{tcolorbox}
\end{center}

@@

df_area_age_behav <- df %>%
  select(iso3, area_id, age_group, population_nosex12m, incidence_nosex12m, starts_with("population_sex"), starts_with("incidence_sex")) %>%
  pivot_longer(
    cols = starts_with("population") | starts_with("incidence"),
    names_to = c("indicator", "category"),
    names_sep = "_"
  ) %>%
  pivot_wider(
    id_cols = c("iso3", "area_id", "age_group", "indicator", "category"),
    names_from = "indicator",
    values_from = "value"
  )
```

```{r echo=FALSE}
df_area_age_behav %>%
  arrange(desc(incidence)) %>%
  head() %>%
  knitr::kable()
```

@@

```{r include=FALSE}
df_area_age <- df_area_age_behav %>%
  group_by(iso3, area_id, age_group) %>%
  summarise(
    incidence = sum(incidence * population, na.rm = TRUE) / sum(population, na.rm = TRUE),
    population = sum(population, na.rm = TRUE)
  ) %>%
  ungroup()
```

```{r echo=FALSE}
df_area_age %>%
  select(iso3, area_id, age_group, population, incidence) %>%
  arrange(desc(incidence)) %>%
  head() %>%
  knitr::kable()
```

@@

@@

* ?

* Infections reached analysis simplistic, doesn't account for e.g. difficulty in reaching each strata, variable intervention effectiveness, change in strata membership
* High risk sexual behaviours are under-reported, and likely varyingly so by age (foremost) and location (less so)
* Risk group definitions didn't consider other important characteristics that may determine risk e.g. condom usage
* Our analysis only focused on AGYW 15-29, and could be extended to adults of both sexes aged 15-49

>>>>

1. Enable implementation of prevention strategies stratified by incidence and behaviour
2. Assess the benefits of such approaches

@@

1. Estimate the proportion of AGYW in four sexual risk behaviour groups at a district level (in 13 countries identified as priority by The Global Fund)
2. Analyze numbers of new infections reached by stratified prevention strategies

@@

* Sexual behaviour data from AIS, BAIS, DHS and PHIA household surveys (see next slide)
* Later, we used district-level HIV incidence, prevalence, population size estimates from the Naomi model [@eaton2021naomi]

@@

### Stage 1: $k = 1, 2, 3^{+}$

* Multinomial logistic regression model for the proportion of AGYW in the none, low and {high, very high} risk groups, using all 47 surveys
* Selected model included:
  * Age effects (IID)
  * Country effects (IID)
  * Correlated spatial effects (ICAR)
  * Correlated temporal effects (AR1) 
* Multinomial-Poisson transformation allowed us to use `R-INLA` for inference

@@

### Stage 2: $k = 3, 4$

* Logistic regression model for the proportion of those in the {high, very high} risk groups who are in the very high risk group, using only the 13 surveys with a specific transactional sex question
* Selected model included:
  * Age effects (IID)
  * Country effects (IID)
  * Correlated spatial effects (ICAR)
  * Clients of FSW covariates [@hodgins2022population]
* Also used `R-INLA` for inference here

@@

* Take 1000 samples from each model, then multiply suitably to generate estimates for all four risk groups
* We adjusted the very high risk group to match national FSW estimates from @johnston2022deriving

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ Estimates of risk group proportions $\rho_{itak}$ by district, year and age group
\end{tcolorbox}
\end{center}

```{r echo=FALSE, out.width = "400px"}
knitr::include_graphics("depends/aids-abstract-A.pdf")
```

```{r echo=FALSE, out.width = "400px"}
knitr::include_graphics("depends/aids-abstract-B.pdf")
```

@@

<!-- ```{r echo=FALSE, out.width = "400px"} -->
<!-- knitr::include_graphics("depends/3p1-within-between-country-variation.png") -->
<!-- ``` -->

@@

### Integration of all relevant surveys

* Modular approach allowed estimating FSW proportion even for surveys without a specific transactional sex question

### Alleivating small-sample sizes

* Direct estimates display large variance
* We borrowed information across space, between countries and over surveys
* Our estimates more plausibly reflect spatial heterogeneity

@@

### Estimates even where there isn't direct data

* Although some people think of this as "making up data", the data almost never "speak for themselves"
* Uncertainty should be higher in regions with infilling: important to clearly communicate this

@@

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ Estimates of HIV incidence $\lambda_{iak}$ and number of new HIV infections $I_{iak}$ by district, age group and risk group
\end{tcolorbox}
\end{center}

@@

df_area_age_behav <- df %>%
  select(area_id, age_group, population_nosex12m, incidence_nosex12m, starts_with("population_sex"), starts_with("incidence_sex")) %>%
  pivot_longer(
    cols = starts_with("population") | starts_with("incidence"),
    names_to = c("indicator", "category"),
    names_sep = "_"
  ) %>%
  pivot_wider(
    id_cols = c("area_id", "age_group", "indicator", "category"),
    names_from = "indicator",
    values_from = "value"
  )
```

```{r echo=FALSE}
df_area_age_behav %>%
  arrange(desc(incidence)) %>%
  head() %>%
  knitr::kable(digits = 2)
```

@@

```{r include=FALSE}
df_area_age <- df_area_age_behav %>%
  group_by(area_id, age_group) %>%
  summarise(
    incidence = sum(incidence * population, na.rm = TRUE) / sum(population, na.rm = TRUE),
    population = sum(population, na.rm = TRUE)
  ) %>%
  ungroup()
```

```{r echo=FALSE}
df_area_age %>%
  select(area_id, age_group, population, incidence) %>%
  arrange(desc(incidence)) %>%
  head() %>%
  knitr::kable(digits = 2)
```

@@

* Simplistic infections reached analysis
  * No accounting for difficulty in reaching each strata
  * Variable intervention effectiveness
  * Change in strata membership
* Under-reporting of high risk sexual behaviours
  * Variation in under-reporting (likely by age, foremost, and location, less so) particularly concerning
* Simple risk groups, not clearly linked to data
  * Didn't consider other important characteristics that may determine risk e.g. condom usage
* Only focused on AGYW 15-29
  * Could be extended to adults of both sexes aged 15-49

@@

* We hope that our risk group estimates will help countries to implement the Global AIDS Strategy
  * (User guide currently under development)
* It's very important to reach FSW!
* Countries have different epidemic profiles
